Single-seismometer Focal Mechanism and Uncertainty Estimation from Body Waves

### 2024-10-29 paper/main.tex
tilde lol

  \subsection{Tolerance derivation}
    Given vectors $\bm{A}_s$ and $\bm{A}_o$, we would like to know the largest cosine
    similarity for which $\bm{A}_s$ is considered an acceptable fit with respect to the recorded amplitudes
    while accounting for the asymmetry of error levels for different amplitudes. Visualized in 3D
    space, we have a confidence ellipsoid centered at $\bm{A}_o$ (equation \ref{eq:7}) such that
    acceptable fits are vectors which fall inside the elliptical cone tangent to it with vertex at the
    origin.
    \begin{align} \label{eq:7}
      E(\bm{p}) = (\bm{A} - \bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = 1
    \end{align}

    \noindent where $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ is arbitrary
    and $\hat{\bm{\Sigma}} = \text{diag}(\sigma_P^2, \sigma_{SV}^2, \sigma_{SH}^2)/\chi^2_{2,\alpha}$ for some $\alpha \in (0,1)$. The elliptical
    cone intersects with the ellipsoid at every boundary point for which
    \begin{align} \label{eq:8}
      (\nabla_{\bm{A}} E)^\top \bm{A} = 2(\bm{A} - \bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} \bm{A} = 0
    \end{align}

    Combining equations \ref{eq:7} and \ref{eq:8} gives the plane containing every point of 
    tangency:
    \begin{align} \label{eq:9}
      \frac{1}{2} (\nabla_{\bm{A}} E)^\top \bm{A} - f(\bm{A}) = 
              \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = -1
    \end{align}

@@

    \begin{align} \label{eq:12}
      \nonumber \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} ((1 - \nicefrac{1}{c_1})\bm{A}_o - \bm{A}_o)  = -1\\
      \Rightarrow c_1 = \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_o = \|\bm{A}_o^\top, \bm{A}_o\|_{\hat{\bm{\Sigma}}^{-1}}^2
    \end{align}
    \begin{align} \label{eq:13}
      \nonumber\bm{D} &= (\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1})^\top \times (\bm{A}_s \times \bm{A}_o)\\
        &= c_1\bm{A}_s - c_2\bm{A}_o
    \end{align}

    \noindent where $c_2 = \bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_s = \langle\bm{A}_o^\top, \bm{A}_s\rangle_{\hat{\bm{\Sigma}}^{-1}}$.

@@

    \begin{align} \label{eq:14}
      \nonumber (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)^\top \hat{\bm{\Sigma}}^{-1} (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)  = 1\\
      ({\bm{D}^\top \hat{\bm{\Sigma}}^{-1} \bm{D}})t^2 - 
          \nicefrac{2}{c_1}(\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{D})t + (\nicefrac{1}{c_1} - 1) = 0
    \end{align}

    Denote $c_3 = \bm{A}_s^\top \hat{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s^\top, \bm{A}_s\|_{\hat{\bm{\Sigma}}^{-1}}^2$. Then by
    substitution we have
    ${\bm{D}^\top \hat{\bm{\Sigma}}^{-1} \bm{D}} = c_1(c_1c_3 - c_2^2)$ and ${\bm{A}_o^\top \hat{\bm{\Sigma}}^{-1} \bm{D}} = 0$.
    Solving then yields the necessary points (denoted $\bm{A}_b$) as
    \begin{align} \label{eq:15}
      \bm{A}_b = \left(1 - \frac{1}{c_1}\right)\bm{A}_o \pm
        \left(\frac{1}{c_1} \sqrt{\frac{c_1 - 1}{c_1c_3 - c_2^2}}\right)(c_1\bm{A}_s - c_2\bm{A}_o)
    \end{align}

    Equation \ref{eq:15} shows that for the points to be well-defined, the following conditions
    must hold relative to the inner-product space defined by $\hat{\bm{\Sigma}}^{-1}$:

    \begin{itemize}
      \item Strict Cauchy-Schwarz inequality: $c_1c_3 > c_2^2$. This also implies that
        $\hat{\bm{\Sigma}} \succ 0$ and $\bm{A}_s \nparallel \bm{A}_o$. Practically, it means that all
        uncertainties are necessary and inversion methods that depend on forward simulation should
        treat parallelism as an edge case.
      \item Confidence ellipsoid cannot contain origin: $c_1 \geq 1$. Practically, we should expect
      errors to be a small fraction of observed amplitudes.
    \end{itemize}

    Thus, given 2 nonparallel vectors $\bm{A}_o$ and $\bm{A}_s$ and positive definite covariance
    matrix $\hat{\bm{\Sigma}}$, we accept the simulation if $\hat{\bm{A}}_o^\top \hat{\bm{A}_s} \geq \text{tol} 
    = \min_b \hat{\bm{A}}_o^\top \hat{\bm{A}_b}$. The condition, more formally, is
    \begin{align} \label{eq:16}
      \max_b \hat{\bm{A}}_o^\top (\hat{\bm{A}}_b - \hat{\bm{A}}_s) \leq 0
    \end{align}

>>>>

  \subsection{Tolerance derivation}
    Given vectors $\bm{A}_s$ and $\bm{A}_o$, we would like to know the largest cosine
    similarity for which $\bm{A}_s$ is considered an acceptable fit with respect to the recorded amplitudes
    while accounting for the asymmetry of error levels for different amplitudes. Visualized in 3D
    space, we have a confidence ellipsoid centered at $\bm{A}_o$ (equation \ref{eq:7}) such that
    acceptable fits are vectors which fall inside the elliptical cone tangent to it with vertex at the
    origin.
    \begin{align} \label{eq:7}
      E(\bm{p}) = (\bm{A} - \bm{A}_o)^\top \tilde{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = 1
    \end{align}

    \noindent where $\bm{A} = (A^P, A^{SV}, A^{SH})^\top$ is arbitrary
    and $\tilde{\bm{\Sigma}} = \text{diag}(\sigma_P^2, \sigma_{SV}^2, \sigma_{SH}^2)/\chi^2_{2,\alpha}$ for some $\alpha \in (0,1)$. The elliptical
    cone intersects with the ellipsoid at every boundary point for which
    \begin{align} \label{eq:8}
      (\nabla_{\bm{A}} E)^\top \bm{A} = 2(\bm{A} - \bm{A}_o)^\top \tilde{\bm{\Sigma}}^{-1} \bm{A} = 0
    \end{align}

    Combining equations \ref{eq:7} and \ref{eq:8} gives the plane containing every point of 
    tangency:
    \begin{align} \label{eq:9}
      \frac{1}{2} (\nabla_{\bm{A}} E)^\top \bm{A} - f(\bm{A}) = 
              \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} (\bm{A} - \bm{A}_o)  = -1
    \end{align}

@@

    \begin{align} \label{eq:12}
      \nonumber \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} ((1 - \nicefrac{1}{c_1})\bm{A}_o - \bm{A}_o)  = -1\\
      \Rightarrow c_1 = \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_o = \|\bm{A}_o^\top, \bm{A}_o\|_{\tilde{\bm{\Sigma}}^{-1}}^2
    \end{align}
    \begin{align} \label{eq:13}
      \nonumber\bm{D} &= (\bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1})^\top \times (\bm{A}_s \times \bm{A}_o)\\
        &= c_1\bm{A}_s - c_2\bm{A}_o
    \end{align}

    \noindent where $c_2 = \bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \langle\bm{A}_o^\top, \bm{A}_s\rangle_{\tilde{\bm{\Sigma}}^{-1}}$.

@@

    \begin{align} \label{eq:14}
      \nonumber (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)^\top \tilde{\bm{\Sigma}}^{-1} (t\bm{D} - \nicefrac{1}{c_1}\bm{A}_o)  = 1\\
      ({\bm{D}^\top \tilde{\bm{\Sigma}}^{-1} \bm{D}})t^2 - 
          \nicefrac{2}{c_1}(\bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{D})t + (\nicefrac{1}{c_1} - 1) = 0
    \end{align}

    Denote $c_3 = \bm{A}_s^\top \tilde{\bm{\Sigma}}^{-1} \bm{A}_s = \|\bm{A}_s^\top, \bm{A}_s\|_{\tilde{\bm{\Sigma}}^{-1}}^2$. Then by
    substitution we have
    ${\bm{D}^\top \tilde{\bm{\Sigma}}^{-1} \bm{D}} = c_1(c_1c_3 - c_2^2)$ and ${\bm{A}_o^\top \tilde{\bm{\Sigma}}^{-1} \bm{D}} = 0$.
    Solving then yields the necessary points (denoted $\bm{A}_b$) as
    \begin{align} \label{eq:15}
      \bm{A}_b = \left(1 - \frac{1}{c_1}\right)\bm{A}_o \pm
        \left(\frac{1}{c_1} \sqrt{\frac{c_1 - 1}{c_1c_3 - c_2^2}}\right)(c_1\bm{A}_s - c_2\bm{A}_o)
    \end{align}

    Equation \ref{eq:15} shows that for the points to be well-defined, the following conditions
    must hold relative to the inner-product space defined by $\tilde{\bm{\Sigma}}^{-1}$:

    \begin{itemize}
      \item Strict Cauchy-Schwarz inequality: $c_1c_3 > c_2^2$. This also implies that
        $\tilde{\bm{\Sigma}} \succ 0$ and $\bm{A}_s \nparallel \bm{A}_o$. Practically, it means that all
        uncertainties are necessary and inversion methods that depend on forward simulation should
        treat parallelism as an edge case.
      \item Confidence ellipsoid cannot contain origin: $c_1 \geq 1$. Practically, we should expect
      errors to be a small fraction of observed amplitudes.
    \end{itemize}

    Thus, given 2 nonparallel vectors $\bm{A}_o$ and $\bm{A}_s$ and positive definite covariance
    matrix $\tilde{\bm{\Sigma}}$, we accept the simulation if $\hat{\bm{A}}_o^\top \hat{\bm{A}_s} \geq \text{tol} 
    = \min_b \hat{\bm{A}}_o^\top \hat{\bm{A}_b}$. The condition, more formally, is
    \begin{align} \label{eq:16}
      \max_b \hat{\bm{A}}_o^\top (\hat{\bm{A}}_b - \hat{\bm{A}}_s) \leq 0
    \end{align}
